visitation_bias

### 2021-06-10 code/ms2_nullmodels.Rmd
finished intro and almost done methods and results

## 2.2 Overall bias by morphotype
```{r}
ggplot(null_tab1, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(14) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Bias")

ggsave("../ms_2/figures/null_model_ktype_preference.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktype_preference.png", height = 6, width = 7)
```

@@

### 2.3.2 Model
```{r}
gam1 <- gam(preference ~ 
              ktype + 
              s(pbl.w, by = ktype, k = 5), 
            data = null_tab1) %>% getViz()

summary(gam1)
check.gamViz(gam1)
#plot.gamViz(gam1, allTerms = TRUE)

# Extract predictions
gam1_pred <- predict_gam(gam1) %>%
  mutate(fit.plus.se = fit + (2*se.fit),
         fit.minus.se = fit - (2*se.fit))

# Plot
ggplot(gam1_pred, aes(pbl.w, fit)) +
  geom_line() + # fit line
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1, # Original data
             aes(pbl.w, preference, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  #scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Bias", 
       color = "Species") +
  theme_light()

ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", height = 6, width = 7)
```

@@

#check.gamViz(gam3.1)
#gam3.1_sum <- summary(gam3.1, re.test = FALSE)
#plot.gamViz(gam3.1, allTerms = TRUE)
```

@@

@@

ggplot(filter(null_tab3_join,
              bb.sp %in% c("hort", "prat", "telu", 
                           "wurf", "soro", "pasc") &
                ktype %in% c("bell", "lip", "flag", "head")), aes(cover, preference)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_grid(bb.sp~ktype, scales = "free") +
  ylim(c(-1, 1))

@@

## 5.1 Overall visitation rates
```{r}
# Get mean proportional visitation by morphotype*bb.sp
visitation_prop <- visitation %>% 
  group_by(sample.id, bb.sp, ktype) %>%
  summarize(visits = n()) %>%
  group_by(sample.id, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  select(sample.id, bb.sp, ktype, prop.visits) %>%
  group_by(bb.sp) %>%
  nest() %>%
  mutate(fill = map(data, function(x) full_join(x, flower_cover_ktype) %>% replace_na(replace = list(prop.visits = 0)))) %>%
  unnest(fill) %>%
  select(-data) %>%
  group_by(bb.sp, ktype) %>%
  summarize(mean.prop.visits = mean(prop.visits)) %>%
  filter(ktype != 10) %>%
   mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels() %>%
  left_join(bb_traits)

null_model_overall <- ggplot(null_tab1, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(10) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Bias")

@@

summary(gam1b)
check.gamViz(gam1b)
#plot.gamViz(gam1, allTerms = TRUE)

@@

gam1b_pred <- predict_gam(gam1b) %>%
  mutate(fit.plus.se = inv_logit(fit + (2*se.fit)),
         fit.minus.se = inv_logit(fit - (2*se.fit)),
         fit = inv_logit(fit))

raw_visits_pbl_GAM <- ggplot(gam1b_pred, aes(pbl.w, fit)) +
  geom_line() + # fit line
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = visitation_prop, # Original data
             aes(pbl.w, mean.prop.visits, color = reorder(bb.sp, pbl.w)), 
             inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Mean proportional visitation", 
       color = "Species") +
  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  theme_light(10)

null_model_pbl_GAM <- ggplot(gam1_pred, aes(pbl.w, fit)) +
  geom_line() + # fit line
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1, # Original data
             aes(pbl.w, preference, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  #scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Bias", 
       color = "Species") +
  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  theme_light(10)

@@

ggsave("../ms_2/figures/ktypeXtonguelength_raw_visitation.pdf", pbl_GAM_plot, height = 5, width = 8.5)
ggsave("../ms_2/figures/ktypeXtonguelength_raw_visitation.png", pbl_GAM_plot, height = 5, width =8.5)
```

@@

# Ordisurf on tongue length
ordisurf_visits <- ordisurf(pca_visits, tongues2$pbl.w) 
summary(ordisurf_visits)

>>>>

## 2.2 Overall bias by morphotype
```{r}
null_model_overall<- ggplot(null_tab1, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(10) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Selectivity")

ggsave("../ms_2/figures/null_model_ktype_preference.pdf", null_model_overall, height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktype_preference.png", null_model_overall, height = 6, width = 7)
```

@@

### 2.3.2 Call model 
```{r}
gam1 <- gam(preference ~ 
              ktype + 
              s(pbl.w, by = ktype, k = 5), 
            data = null_tab1) %>% getViz()

gam1_sum <- summary(gam1)
check.gamViz(gam1)
#plot.gamViz(gam1, allTerms = TRUE)
```

### 2.3.3 Plot
```{r}
# Tabulate model summary
gam1_s_table <- gam1_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  mutate(ktype = str_extract(term, "(?<=ktype).*$")) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype))

# Extract predictions
gam1_pred <- predict_gam(gam1) %>%
  mutate(fit.plus.se = fit + (2*se.fit),
         fit.minus.se = fit - (2*se.fit)) %>%
  left_join(gam1_s_table) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE)) %>%
  mutate(sig = factor(sig, levels = c(TRUE, FALSE)))

# Plot
null_model_pbl_GAM <- ggplot(gam1_pred, aes(pbl.w, fit)) +
  geom_line(aes(linetype = sig)) + # fit line, linetype for significance
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1, # Original data
             aes(pbl.w, preference, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Selectivity", 
       color = "Species", linetype = "p < 0.05") +
  theme_light(10)

ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.pdf", null_model_pbl_GAM, height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", null_model_pbl_GAM, height = 6, width = 7)
```

@@

check.gamViz(gam3.1)
gam3.1_sum <- summary(gam3.1, re.test = FALSE)
plot.gamViz(gam3.1, allTerms = TRUE)

gam3.1b <- bam(preference ~
                bb.ktype +
                s(site, by = bb.ktype, bs = "re") +
                s(year, by = bb.ktype, bs = "re") +
                te(yday, elev.mean, by = bb.ktype),
              family = "gaussian",
              method = "fREML",
              discrete = TRUE,
              data = filter(null_tab3_join, 
                            bb.sp %in% c("hort", "prat", "telu", 
                                         "wurf", "soro", "pasc") &
                              ktype %in% c("bell", "lip", "flag", "head") &
                              prop.cover < 0.5)) %>% 
                getViz()

check.gamViz(gam3.1b)
gam3.1b_sum <- summary(gam3.1b, re.test = FALSE)
plot.gamViz(gam3.1b, allTerms = TRUE)
```

@@

ggsave("../ms_2/figures/preference_lability_gam.pdf", height = 5, width = 6.5)
ggsave("../ms_2/figures/preference_lability_gam.png", height = 5, width = 6.5)

# And now having dropped the high abundance outliers
gam3.1b_s_table <- gam3.1b_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{6}$")) %>%
  separate(term, c("bb.sp", "ktype")) %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"
  )) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(bb.sp, ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.
hort_3_gam3.1b <- plot(sm(gam3.1b, 49))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("bell", n()))

hort_5_gam3.1b <- plot(sm(gam3.1b, 50))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("lip", n()))

hort_6_gam3.1b <- plot(sm(gam3.1b, 51))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("flag", n()))

hort_7_gam3.1b <- plot(sm(gam3.1b, 52))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("head", n()))

pasc_3_gam3.1b <- plot(sm(gam3.1b, 53))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("bell", n()))

pasc_5_gam3.1b <- plot(sm(gam3.1b, 54))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("lip", n()))

pasc_6_gam3.1b <- plot(sm(gam3.1b, 55))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("flag", n()))

pasc_7_gam3.1b <- plot(sm(gam3.1b, 56))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("head", n()))

prat_3_gam3.1b <- plot(sm(gam3.1b, 57))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("bell", n()))

prat_5_gam3.1b <- plot(sm(gam3.1b, 58))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("lip", n()))

prat_6_gam3.1b <- plot(sm(gam3.1b, 59))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("flag", n()))

prat_7_gam3.1b <- plot(sm(gam3.1b, 60))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("head", n()))

soro_3_gam3.1b <- plot(sm(gam3.1b, 61))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("bell", n()))

soro_5_gam3.1b <- plot(sm(gam3.1b, 62))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("lip", n()))

soro_6_gam3.1b <- plot(sm(gam3.1b, 63))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("flag", n()))

soro_7_gam3.1b <- plot(sm(gam3.1b, 64))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("head", n()))

telu_3_gam3.1b <- plot(sm(gam3.1b, 65))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("bell", n()))

telu_5_gam3.1b <- plot(sm(gam3.1b, 66))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("lip", n()))

telu_6_gam3.1b <- plot(sm(gam3.1b, 67))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("flag", n()))

telu_7_gam3.1b <- plot(sm(gam3.1b, 68))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("head", n()))

wurf_3_gam3.1b <- plot(sm(gam3.1b, 69))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("bell", n()))

wurf_5_gam3.1b <- plot(sm(gam3.1b, 70))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("lip", n()))

wurf_6_gam3.1b <- plot(sm(gam3.1b, 71))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("flag", n()))

wurf_7_gam3.1b <- plot(sm(gam3.1b, 72))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("head", n()))

plotting_data__gam3.1b <- bind_rows(hort_3_gam3.1b, hort_5_gam3.1b, hort_6_gam3.1b, hort_7_gam3.1b,
                           pasc_3_gam3.1b, pasc_5_gam3.1b, pasc_6_gam3.1b, pasc_7_gam3.1b,
                           prat_3_gam3.1b, prat_5_gam3.1b, prat_6_gam3.1b, prat_7_gam3.1b,
                           soro_3_gam3.1b, soro_5_gam3.1b, soro_6_gam3.1b, soro_7_gam3.1b,
                           telu_3_gam3.1b, telu_5_gam3.1b, telu_6_gam3.1b, telu_7_gam3.1b,
                           wurf_3_gam3.1b, wurf_5_gam3.1b, wurf_6_gam3.1b, wurf_7_gam3.1b) %>%
  group_by(bb.sp, ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(bb_traits) %>%
  left_join(gam3.1b_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         bb.sp, ktype, pbl.w, edf, pval) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))

ggplot(plotting_data__gam3.1b, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40,
               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
  scale_linetype_manual(values = c("dotted", "solid"))+
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

@@

ggplot(filter(null_tab3_join,
              bb.sp %in% c("hort", "prat", "telu", 
                           "wurf", "soro", "pasc") &
                ktype %in% c("bell", "lip", "flag", "head") &
                prop.cover < 0.5), aes(prop.cover, preference)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_grid(bb.sp~ktype) +
  ylim(c(-1, 1))

ggplot(filter(abundance, ktype %in% c("bell", "lip", "flag", "head")), 
       aes(prop.cover)) +
  geom_histogram() +
  facet_wrap(~ktype)

ggplot(abundance, aes(prop.cover)) +
  geom_histogram() +
  facet_wrap(~ktype)

@@

## 5.1 Overall visitation rates
```{r}
# Get mean proportional visitation by morphotype*bb.sp
visitation_prop <- visitation %>% 
  group_by(sample.id, bb.sp, ktype) %>%
  summarize(visits = n()) %>%
  group_by(sample.id, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  select(sample.id, bb.sp, ktype, prop.visits) %>%
  group_by(bb.sp) %>%
  nest() %>%
  mutate(fill = map(data, function(x) full_join(x, flower_cover_ktype) %>% replace_na(replace = list(prop.visits = 0)))) %>%
  unnest(fill) %>%
  select(-data) %>%
  group_by(bb.sp, ktype) %>%
  summarize(mean.prop.visits = mean(prop.visits)) %>%
  filter(ktype != 10) %>%
   mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels() %>%
  left_join(bb_traits) %>%
  ungroup() %>%
  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w))

# null_model_overall <- ggplot(null_tab1, aes(ktype, preference)) +
#   #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
#   geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
#   geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
#   theme_light(10) +
#   scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
#   labs(x = "Floral morphotype",
#        y = "Bias")

@@

gam1b_sum <- summary(gam1b)
check.gamViz(gam1b)
#plot.gamViz(gam1, allTerms = TRUE)

@@

# Tabulate model summary
gam1b_s_table <- gam1b_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  mutate(ktype = str_extract(term, "(?<=ktype).*$")) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype))

# Extract predictions
gam1b_pred <- predict_gam(gam1b) %>%
  mutate(fit.plus.se = inv_logit(fit + (2*se.fit)),
         fit.minus.se = inv_logit(fit - (2*se.fit)),
         fit = inv_logit(fit)) %>%
  left_join(gam1b_s_table) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE)) %>%
  mutate(sig = factor(sig, levels = c(TRUE, FALSE)))

# Plot
raw_visits_pbl_GAM <- ggplot(gam1b_pred, aes(pbl.w, fit)) +
  geom_line(aes(linetype = sig)) + # fit line, linetype for significance
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = visitation_prop, # Original data
             aes(pbl.w, mean.prop.visits, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Mean proportional visitation", 
       color = "Species", linetype = "p < 0.05") +
  theme_light(10)

@@

ggsave("../ms_2/figures/ktypeXtonguelength_paired.pdf", pbl_GAM_plot, height = 5, width = 9.5)
ggsave("../ms_2/figures/ktypeXtonguelength_paired.png", pbl_GAM_plot, height = 5, width = 9.5)
```

@@

# Ordisurf on tongue length
ordisurf_visits <- ordisurf(pcoa_visits, tongues2$pbl.w) 
summary(ordisurf_visits)
